---
title: "Homework 2"
author: "Ruwen Zhou"
date: "9/28/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

```{r}
path_to_data = "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx"
```


Read the Mr. Trashwheel dataset. 

```{r}
trashwheel_df = 
	read_xlsx(
		path = path_to_data,
		sheet = "Mr. Trash Wheel",
		range = cell_cols("A:N")) %>% 
	janitor::clean_names() %>% 
	drop_na(dumpster) %>% 
	mutate(
		sports_balls = round(sports_balls),
		sports_balls = as.integer(sports_balls)
	)
```

```{r}
trashwheel_df
```

Read precipitation data! For 2018 and 2017. 

```{r}
df_2018 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2018 Precipitation",
		skip = 1) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2018) %>% 
	relocate(year)
  #mutate(month = month.name)

df_2017 = 
	read_excel(
		"./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
		sheet = "2017 Precipitation",
		skip = 1) %>% 
	janitor::clean_names() %>% 
	drop_na(month) %>% 
	mutate(year = 2017) %>% 
	relocate(year)
  #mutate(month = month.name) 
```





```{r}
df_2017
df_2018
```


```{r}
month_df = 
	tibble(
		month = 1:12,
		month_name = month.name
	)

precip_df = 
	bind_rows(df_2018, df_2017)
precip_df =
	left_join(precip_df, month_df, by = "month")
```

```{r}
month_df
precip_df
```

## write something about the data set
The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data. In this dataset:

* The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
* The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.


# Problem 2

## load and clean data
```{r}

rm(list = ls())

ny_df = 
  read_csv(
    "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada)  %>%
  mutate(entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)) 
```
```{r}
ny_df
```
## write something about the data set
This dataset is NYC transit data which contains information related to subway stations.It contains variables: line, station name, latitude, longitude, entry, entrance_type, vending, ada, and route1 to route11:

* This dataset has a total number of `r nrow(ny_df)` observations and `r ncol(ny_df)` columns.
* We clean the columns and recode 'entry' from a character variable into a logical variable.
* This dataset is not tidy, especially for rounte data.

## Tidy data

```{r}
ny_tidy = 
  ny_df %>% 
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name", 
    values_to = "route_number") %>% 
  drop_na(route_number)
```

```{r}
ny_df %>% distinct(station_name, line) %>% nrow
ny_df %>% distinct(station_name, line, ada) %>% filter(ada == TRUE) %>% nrow
(ny_tidy %>% filter(vending == 'NO') %>% summarise(sum(entry) / n()))
```


* There are `r ny_df %>% distinct(station_name, line) %>% nrow` distinct stations
* `r ny_df %>% distinct(station_name, line, ada) %>% filter(ada == TRUE) %>% nrow ` are ADA compliant
* `r (ny_tidy %>% filter(vending == 'NO') %>% summarise(sum(entry) / n()))*100` percent of station entrances/exits without vending allow entrance.


# Problem 3

## first question

```{r}
pols_df=
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate_at(vars(month),as.numeric) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president=ifelse(prez_gop==1,"gop","dem")) %>% 
  select(year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem, president)
  
  
```
```{r}
pols_df
```
## Second question

```{r}
snp_df=
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, into = c("day", "month", "year"), sep = "/") %>% 
  mutate_at(vars(month),as.numeric) %>% 
  mutate(month = month.name[month]) %>%  
  arrange(year,month) %>% 
  relocate(year,month)
```


```{r}
snp_df
```

## Third 

```{r}
unemployment_df=
  read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to="unemploy_rate") %>% 
  mutate(month = apply(as.matrix(month),1,
                       function(x) month.name[[
                         match(x,str_to_lower(month.abb))]])) %>% 
  mutate_at(vars(year),as.character)
  

```


```{r}
unemployment_df
```


## join datasets
```{r}
join_df = pols_df %>% 
  left_join(snp_df,by = c("year","month")) %>% 
  left_join(unemployment_df, by = c("year","month"))
```

```{r}
join_df
```

## write something about the dataset
* In pols_df, ranging from year `r min(pols_df$year)` to ``r max(pols_df$year)`. There are `r nrow(pols_df)` politicians and `r ncol(pols_df)` variables related to the number of national politicians who are democratic or republican at any given time: including prez_gop, gov_gop,	sen_gop,	rep_gop,	prez_dem,	gov_dem,	sen_dem,	rep_dem.
* In snp_df, ranging from year `r min(snp_df$year)` to ``r max(snp_df$year)`. There are `r nrow(snp_df)` stocks and `r ncol(snp_df)` variables: "date" is the date of observation and "close" is the closed price of the stock on the date.
* In unemployment_df, ranging from year `r min(unemployment_df$year)` to ``r max(unemployment_df$year)`. There are `r nrow(unemployment_df)` years and `r ncol(unemployment_df)` variables. The variables represent the percentage of unemployment in the month of the the year.
* In join_df, I put all data into one dataset with left join, which makes empty variables for each observations to be NA. There are `r nrow(join_df)` observations and `r ncol(join_df)` variables, ranging from year `r min(join_df$year)` to ``r max(join_df$year)`.

























